State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-05-06 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
Field estimation plot added to the paper

The sensors were spaced in a regular square grid, where the number of sensors was 100, sensor width $\sigma^2_m$ was 1.44 which equals $2$~mm full width at half maximum, with spacing $\Delta_y$ equal to $2.22$~mm. This arrangement modelled the recording as having some cross talk, which is typical from neurophysiological recordings. Note, sensor design to guarantee representation of the field can be achieved using the frequency analysis from section~\ref{SpectralAnalysisSection} given access to the true field. \omg{From frequency analysis the separation is $2.77$, with $\rho = 2$ (note same as basis function).} 
   \begin{figure}\label{fig:experimental design}
   	\begin{center}
   		\includegraphics[width=0.3\textwidth]{./Graph/ExperimentFigurePy_1.pdf} 
   	\end{center}
   	\caption{Example of experimental design and model output. \textbf{a}. Example of the neural field with sensors. The centres of the sensors are shown by crossed and the sensor widths ($1/2$ height of Gaussian) are shown by the circles. \textbf{b} Example of data generated by the model through the observations. \textbf{c}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.} 
   \end{figure}
% \begin{figure}[th]\label{fig:experimental design}
% \begin{center}
% \subfigure[][]{\includegraphics[width=0.35\textwidth]{./Graph/ExperimentFigurePy.pdf}}\\
% \subfigure[][]{\includegraphics[width=0.35\textwidth]{./Graph/ExperimentFigureEEGPy.pdf}}\\
% \subfigure[][]{\includegraphics[width=0.35\textwidth]{./Graph/ExperimentFigurePSDPy.pdf}}
% \end{center}
% \caption{Example of experimental design and model output. \textbf{a}. Example of the neural field with sensors. The centres of the sensors are shown by crossed and the sensor widths ($1/2$ height of Gaussian) are shown by the circles. \textbf{b} Example of data generated by the model through the observations. \textbf{c}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.} 
% \end{figure}
\begin{figure}[th]
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/zeta.pdf}}
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta0.pdf}}\\
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta1.pdf}}
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta2.pdf}}
\caption{Distribution of the parameters estimates over 272
runs.}
\label{fig:Parameters}
\end{figure}
\subsection{Spatial Frequency Analysis} In accordance with Section~\ref{SpectralAnalysisSection}, the spatial frequency of the field was used to confirm the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{} shows the spatial frequency of the field. The cut-off frequency, $\nu_c$, was taken to be \omg{??}, which corresponds to the \omg{??-3~dB} from the peak power. From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of \omg{??}. This confirms our separation was enough to prevent problems associated with aliasing.

@@

\subsection{State and Parameter Estimates} To demonstrate the performance of the parameter and state estimation, $200$ realisations of the data were generated. Each realisation consisted of 200 ms of data and the estimation was applied to the final 200 ms, allowing the model to stabilise. The initial state and parameters were unknown to the estimator. The state estimates were evaluated by comparing the field reconstruction to the true field using the mean over time of the RMSE and standard deviation of the RMSE. The distribution of the parameters and the resulting kernel are shown in figure~\ref{fig:Parameters} and figure~\ref{fig:KernelEstimates} respectively. Although a wide distribution can be seen over the parameters the ratio between them remains almost constant as shown in figure~\ref{}.

@@

>>>>

The sensors were spaced in a regular square grid, where the number of sensors was 100, sensor width $\sigma^2_m$ was 1.44 which equals $2$~mm full width at half maximum, with spacing $\Delta_y$ equal to $2.22$~mm. This arrangement modelled the recording as having some cross talk, which is typical from neurophysiological recordings. Note, sensor design to guarantee representation of the field can be achieved using the frequency analysis from section~\ref{SpectralAnalysisSection} given access to the true field. \omg{From frequency analysis the separation is $2.77$, with $\rho = 2$ (note same as basis function).} 
   \begin{figure}[th]\label{fig:experimental design}
   	\begin{center}
   		\includegraphics[width=0.3\textwidth]{./Graph/ExperimentFigurePy_1.pdf} 
   	\end{center}
   	\caption{Example of experimental design and model output. \textbf{a}. Example of the neural field with sensors. The centres of the sensors are shown by crossed and the sensor widths ($1/2$ height of Gaussian) are shown by the circles. \textbf{b} Example of data generated by the model through the observations. \textbf{c}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.} 
   \end{figure}
% \begin{figure}[th]\label{fig:experimental design}
% \begin{center}
% \subfigure[][]{\includegraphics[width=0.35\textwidth]{./Graph/ExperimentFigurePy.pdf}}\\
% \subfigure[][]{\includegraphics[width=0.35\textwidth]{./Graph/ExperimentFigureEEGPy.pdf}}\\
% \subfigure[][]{\includegraphics[width=0.35\textwidth]{./Graph/ExperimentFigurePSDPy.pdf}}
% \end{center}
% \caption{Example of experimental design and model output. \textbf{a}. Example of the neural field with sensors. The centres of the sensors are shown by crossed and the sensor widths ($1/2$ height of Gaussian) are shown by the circles. \textbf{b} Example of data generated by the model through the observations. \textbf{c}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.} 
% \end{figure}
\begin{figure}[th]
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/zeta.pdf}}
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta0.pdf}}\\
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta1.pdf}}
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta2.pdf}}
\caption{Distribution of the parameters estimates over 272
runs.}
\label{fig:Parameters}
\end{figure}
\subsection{Spatial Frequency Analysis} In accordance with Section~\ref{SpectralAnalysisSection}, the spatial frequency of the field was used to confirm the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{} shows the spatial frequency of the field. The cut-off frequency, $\nu_c$, was taken to be \omg{??}, which corresponds to the \omg{??-3~dB} from the peak power. From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of \omg{??}. This confirms our separation was enough to prevent problems associated with aliasing.

@@

\subsection{State and Parameter Estimates} To demonstrate the performance of the parameter and state estimation, $200$ realisations of the data were generated. Each realisation consisted of 200 ms of data and the estimation was applied to the final 200 ms, allowing the model to stabilise. The initial state and parameters were unknown to the estimator. The state estimates were evaluated by comparing the field reconstruction to the true field using the mean over time of the RMSE and standard deviation of the RMSE. The distribution of the parameters and the resulting kernel are shown in figure~\ref{fig:Parameters} and figure~\ref{fig:KernelEstimates} respectively. Although a wide distribution can be seen over the parameter estimates the ratio between them remans in a narrow range as shown in figure~\ref{fig:ParametersRatio}. One snapshot for the original and the field estimate and the corresponding error are illustrated in figure~\ref{fig:FieldEstimate}.

@@

\begin{figure*}[!th] 
\subfigure[][]{\includegraphics[width=2in]{./Graph/TrueField90.pdf}}
\subfigure[][]{\includegraphics[width=2in]{./Graph/EstimatedField90.pdf}}
\subfigure[][]{\includegraphics[width=2in]{./Graph/EstimationError90.pdf}}\\
\caption{One snapshot of the spatial field; a: True field; b: Estimated field; c: The absolute error.}
\label{fig:FieldEstimate}
\end{figure*}
